WASP-4\lowercase{b} Arrived Early for the TESS Mission

### 2018-12-18 paper/ms.tex
update author list

\author[0000-0002-5720-4827]{M. Hill}

@@

\author[0000-0002-7084-0529]{S. Kane}

@@

@@

\author{G. R. Ricker} % grr@space.mit.edu WAITING

@@

\author[0000-0002-5322-2315]{A. Glidden} % aglidden@mit.edu WAITING

@@

WASP-4 was observed by TESS with Camera 2 from August 23 to September
20, 2018, within the second ``sector'' of science operations.  The
star is designated as TIC 402026209 in the TESS Input Catalog
\citep{stassun_TIC_2018}.  The pixel data for an $11\times11$ array
surrounding WASP-4 were averaged into 2-minute stacks by the onboard
computer.  The data were downlinked via the Deep Space Network. The
timestamps from the spacecraft clock were then transformed by the
Payload Operations Center into the {\it Temps Dynamique Barycentrique}
(TDB) reference system.  The images were then reduced to lightcurves
by the Science Processing Operations Center (SPOC) at NASA
Ames~\citep{jenkins_tess_2016}.  During this processing, the SPOC used
the known spacecraft trajectory to compute the barycentric time
corrections on a target by target basis, and expressed the timestamps
as Barycentric TESS Julian Dates (BTJD), which is simply the
barycentric Julian date minus 2{,}457{,}000.  The lightcurves were
vetted and released by the MIT TESS Science Office to the Mikulski
Archive for Space Telescopes on November 29,
2018~\citep{ricker_tess_alerts_2018}.

@@

\acknowledgements
L.G.B.\ gladly acknowledges helpful discussions with
A.~Bailey, F.~Dai and V.~Van Eylen, and is grateful to the
people who have turned TESS from an idea into reality.
%
J.N.W.\ thanks the TESS project and the Heising-Simons foundation for
supporting this work.
%
WASP-4 was included on the ``short-cadence'' target list thanks to the
Guest Investigator programs of J.\ Southworth and S.\ Kane (G011112
and G011183 respectively). 
%
This paper includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
This research has made use of the NASA Exoplanet Archive, which is
operated by the California Institute of Technology, under contract
with the National Aeronautics and Space Administration under the
Exoplanet Exploration Program.
%
This work made use of NASA's Astrophysics Data System Bibliographic
Services.
%
This research has made use of the VizieR catalogue access tool, CDS,
Strasbourg, France. The original description of the VizieR service was
published in A\&AS 143, 23.
%
This work has made use of data from the European Space Agency (ESA)
mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
for the DPAC has been provided by national institutions, in particular
the institutions participating in the {\it Gaia} Multilateral
Agreement.
%
\newline
%
\facility{
	TESS \citep{ricker_transiting_2015},
	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018}
}
%
\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  \texttt{astropy} \citep{the_astropy_collaboration_astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  \texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{corner} \citep{corner_2016},
  \texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{IPython} \citep{perez_2007},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001}.
}

@@

@@

The TESS lightcurve files provide observation start and end times in
three different time systems: ${\rm JD}_{\rm UTC}$, ${\rm JD}_{\rm
TDB}$, and ${\rm BJD}_{\rm TDB}$.  First, we verified for a few select
lightcurve files that ${\rm JD}_{\rm UTC}$ lagged behind ${\rm
JD}_{\rm TDB}$ by the expected $32.184 + N\,{\rm seconds}$, where $N$
is the number of leap-seconds since 1961. For the relevant observation
time, $N=37$, and the offset was as expected.  Then, using the ${\rm
JD}_{\rm UTC}$ timestamp, we calculated the barycentric correction
using the software written by \citealt{eastman_achieving_2010}.  (For
simplicity, we assumed the TESS spacecraft was located at the Earth's
geocenter, because the light-travel time between the geocenter and the
spacecraft is on the order of only one second.)  This gave us times in
${\rm BJD}_{\rm TDB}$ that agreed with the ${\rm BJD}_{\rm TDB}$ times
that are delivered with the TESS data products.

@@

\begin{figure*}[ht!]
  \begin{center}
    \leavevmode
    \includegraphics[width=0.9\textwidth]{f7.pdf}
  \end{center}
  \vspace{-0.5cm}
  \caption{
    {\bf There is no evidence for a systematic offset between TESS
    times and the barycentric reference.} While the WASP-4b transits
    fell about 82 seconds earlier than expected, other well-observed
    hot Jupiters, in particular WASP-6b and WASP-18b, arrived on time.
    Ticks are observed TESS transit midtimes; the orange distribution
    is a gaussian centered on zero with standard deviation
    ($\sigma_{\rm predicted}$) calculated from the pre-TESS transit
    times.  The blue distribution is a gaussian centered on the
    weighted average of the TESS times, with width equal to the
    uncertainty in the mean, {\it i.e.}, the standard deviation
    of the TESS residual times divided by $\sqrt{N-1}$, with $N$ the
    number of transits.
    \label{fig:hjs}
  }
\end{figure*}

@@

To use these results to place a quantitative limit on any global clock
offset, for each hot Jupiter we considered the model
\begin{equation}
  t_{\rm tra}(E) = t_0 + PE + t_{\rm offset},
\end{equation}
for $t_{\rm offset}$ a systematic constant offset between the reported
timestamps and the true ${\rm BJD}_{\rm TDB}$ reference.  Our priors
were
\begin{align}
  t_0 &\sim \mathcal{N}[t_0', \sigma_{t_0'}], \\
  P &\sim \mathcal{N}[P', \sigma_{P'}], \\
  t_{\rm offset} &\sim \mathcal{U}[-20\sigma_{t_0'},20\sigma_{t_0'}],
\end{align}
where $\mathcal{N}$ and $\mathcal{U}$ denote a normal and uniform
distribution, $(t_0', P')$ are the best-fit reference time and period
using only the pre-TESS transit times, and $(\sigma_{t_0'},
\sigma_{P'})$ are the corresponding uncertainties.

@@

% \section{Verifying the WASP-4 Archival Times}
% \label{sec:verify_archival_times}
% 
% \begin{figure*}[ht!]
%   \begin{center}
%     \leavevmode
%     \includegraphics[width=0.9\textwidth]{f7.pdf}
%   \end{center}
%   \caption{
%     {\bf Extra times for WASP-4b neither confirm nor refute the
%     	period-decay.}
%     Our timing analysis (Section~\ref{sec:timing}) imposes stringent
%     selection criteria for lightcurves.
%     Here we overplot additional points from the Exoplanet Transit
%     Database (ETD), and the discovery epoch from \citet{wilson_wasp-4b_2008}.
%     \label{fig:verify}
%   }
% \end{figure*}
% 
% The preceding analysis shows that the TESS timestamps are consistent
% at the requisite level with the ${\rm BJD}_{\rm TDB}$ reference.
% Another explanation for the WASP-4b timing variation could be
% incorrect archival times.
% As a sanity check on the times we included for our analysis in
% Table~1, we collected times from the Exoplanet Transit Database
% \citep[ETD; ][]{poddany_ETD_2010}.  We considered only the 12
% lightcurves of the highest self-reported quality (${\rm DQ} = 1$).  We
% then inspected each lightcurve by eye.  If the lightcurve showed
% substantial red noise, we discarded it.  This left 9 midtimes,
% reported in ${\rm HJD}_{\rm UTC}$. We converted the times to ${\rm
% BJD}_{\rm TDB}$ \citep{eastman_achieving_2010}.
% 
% The resulting times are shown in Figure \ref{fig:verify}.  Focusing on
% the epochs that overlap with \citet{huitson_gemini_2017}'s study, four
% of six possible times agree with the \citet{huitson_gemini_2017}
% times; the other two fall about 90 seconds early.  With their stated
% uncertainties, these 6 ETD data points are inconsistent with a
% constant period.  The 6 points came from 5 different observers.
% Systematic errors in absolute clock offsets by different observers
% should be uncorrelated; if there were such systematic errors in the
% ETD times, this could explain the many-sigma outliers.  Though the
% majority of the reported times do support the accuracy of the
% \citet{huitson_gemini_2017} times, a systematic procedure would call
% for averaging the ETD times.  If we did this, and discarded the four
% \citet{huitson_gemini_2017} epochs, it is likely that the evidence for
% period decay would become much weaker.  Since
% \citet{huitson_gemini_2017} documented their time system clearly, and
% produced lightcurves of extremely high quality, we included their times
% in our analysis.  From private correspondence with
% \citealt{huitson_gemini_2017}, the authors also stand by their
% timestamps. Since the ETD data come from
% heterogeneous sources, their timestamps are less clearly documented,
% and the times are thus more prone to systematic errors, we
% categorically omit them from consideration.
% 
% We also overplotted the discovery epoch reported by
% \citet{wilson_wasp-4b_2008}. 
% This epoch folded together WASP-S survey data, a partial FTS
% lightcurve, and a complete EulerCam lightcurve. 
% Though not explicitly stated by \citet{wilson_wasp-4b_2008}, the
% archival WASP-S lightcurves are in ${\rm HJD}_{\rm UTC}$, so they can
% be converted to ${\rm BJD}_{\rm TDB}$ without uncertainty in the
% absolute time reference (Collier-Cameron, priv.\ comm.).
% The discovery epoch falls about 2 minutes earlier than the epochs we
% used for the fit, and is visible in the lower left of Figure
% \ref{fig:verify}.

>>>>

\author[0000-0002-5720-4827]{M. L. Hill}

@@

\author[0000-0002-7084-0529]{S. R. Kane}

@@

\author{J. Villasenor}

@@

\author{G. R. Ricker} % grr@space.mit.edu CONFIRMED

@@

\author[0000-0002-5322-2315]{Ana Glidden} % aglidden@mit.edu CONFIRMED

@@

WASP-4 was observed by TESS with Camera 2 from August 23 to September
20, 2018, within the second ``sector'' of science operations.  The
star is designated as TIC 402026209 in the TESS Input Catalog
\citep{stassun_TIC_2018}.  The pixel data for an $11\times11$ array
surrounding WASP-4 were averaged into 2-minute stacks by the onboard
computer.  The data were downlinked via the Deep Space Network, and the
spacecraft time stamps were calibrated against the ground-station clocks.
The spacecraft clock times were then transformed by the
Payload Operations Center into the {\it Temps Dynamique Barycentrique}
(TDB) reference system.  The images were then reduced to lightcurves
by the Science Processing Operations Center (SPOC) at NASA
Ames~\citep{jenkins_tess_2016}.  During this processing, the SPOC used
the known spacecraft trajectory to compute the barycentric time
corrections on a target by target basis, and expressed the timestamps
as Barycentric TESS Julian Dates (BTJD), which is simply the
barycentric Julian date minus 2{,}457{,}000.  The lightcurves were
vetted and released by the MIT TESS Science Office to the Mikulski
Archive for Space Telescopes on November 29,
2018~\citep{ricker_tess_alerts_2018}.

@@

\acknowledgements
L.G.B.\ and J.N.W.\ gladly acknowledge helpful discussions with
A.~Bailey, F.~Dai, K.~Patra and V.~Van Eylen, and are grateful to the
people who have turned TESS from an idea into reality.
%
J.N.W.\ thanks the TESS project and the Heising-Simons foundation for
supporting this work.
%
WASP-4 was included on the ``short-cadence'' target list thanks to the
Guest Investigator programs of J.\ Southworth and S.\ Kane (G011112
and G011183 respectively). 
%
This paper includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
This research has made use of the NASA Exoplanet Archive, which is
operated by the California Institute of Technology, under contract
with the National Aeronautics and Space Administration under the
Exoplanet Exploration Program.
%
This work made use of NASA's Astrophysics Data System Bibliographic
Services.
%
This research has made use of the VizieR catalogue access tool, CDS,
Strasbourg, France. The original description of the VizieR service was
published in A\&AS 143, 23.
%
This work has made use of data from the European Space Agency (ESA)
mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed
by the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
for the DPAC has been provided by national institutions, in particular
the institutions participating in the {\it Gaia} Multilateral
Agreement.
%
\newline
%
\facility{
	TESS \citep{ricker_transiting_2015},
	Gaia \citep{gaia_collaboration_gaia_2016,gaia_collaboration_gaia_2018}
}
%
\software{
  \texttt{astrobase} \citep{bhatti_astrobase_2018},
  \texttt{astropy} \citep{the_astropy_collaboration_astropy_2018},
  \texttt{astroquery} \citep{astroquery_2018},
  \texttt{BATMAN} \citep{kreidberg_batman_2015},
  \texttt{corner} \citep{corner_2016},
  \texttt{emcee} \citep{foreman-mackey_emcee_2013},
  \texttt{IPython} \citep{perez_2007},
  \texttt{matplotlib} \citep{hunter_matplotlib_2007}, 
  \texttt{numpy} \citep{walt_numpy_2011}, 
  \texttt{pandas} \citep{mckinney-proc-scipy-2010},
  \texttt{scikit-learn} \citep{scikit-learn},
  \texttt{scipy} \citep{jones_scipy_2001}.
}

@@

\begin{figure*}[t!]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f7.pdf}
	\end{center}
	\vspace{-0.5cm}
	\caption{
		{\bf There is no evidence for a systematic offset between TESS
			times and the barycentric reference.} While the WASP-4b transits
		fell about 82 seconds earlier than expected, other well-observed
		hot Jupiters, in particular WASP-6b and WASP-18b, arrived on time.
		Ticks are observed TESS transit midtimes; the orange distribution
		is a gaussian centered on zero with standard deviation
		($\sigma_{\rm predicted}$) calculated from the pre-TESS transit
		times.  The blue distribution is a gaussian centered on the
		weighted average of the TESS times, with width equal to the
		uncertainty in the mean, {\it i.e.}, the standard deviation
		of the TESS residual times divided by $\sqrt{N-1}$, with $N$ the
		number of transits.
		\label{fig:hjs}
	}
\end{figure*}

@@

The TESS lightcurve files provide observation start and end times in
three different time systems: ${\rm JD}_{\rm UTC}$, ${\rm JD}_{\rm
TDB}$, and ${\rm BJD}_{\rm TDB}$.  First, we verified for a few select
lightcurve files that ${\rm JD}_{\rm UTC}$ lagged behind ${\rm
JD}_{\rm TDB}$ by the expected $32.184 + N\,{\rm seconds}$, where $N$
is the number of leap-seconds since 1961. For the relevant observation
time, $N=37$, and the offset was as expected.  Then, using the ${\rm
JD}_{\rm UTC}$ time stamp, we calculated the barycentric correction
using the software written by \citealt{eastman_achieving_2010}.  (For
simplicity, we assumed the TESS spacecraft was located at the Earth's
geocenter, because the light-travel time between the geocenter and the
spacecraft is on the order of only one second.)  This gave us times in
${\rm BJD}_{\rm TDB}$ that agreed with the ${\rm BJD}_{\rm TDB}$ times
that are delivered with the TESS data products.

@@

@@

To use these results to place a quantitative limit on any global clock
offset, for each hot Jupiter we considered the model
\begin{equation}
  t_{\rm tra}(E) = t_0 + PE + t_{\rm offset},
\end{equation}
for $t_{\rm offset}$ a systematic constant offset between the reported
time stamps and the true ${\rm BJD}_{\rm TDB}$ reference.  Our priors
were
\begin{align}
  t_0 &\sim \mathcal{N}[t_0', \sigma_{t_0'}], \\
  P &\sim \mathcal{N}[P', \sigma_{P'}], \\
  t_{\rm offset} &\sim \mathcal{U}[-20\sigma_{t_0'},20\sigma_{t_0'}],
\end{align}
where $\mathcal{N}$ and $\mathcal{U}$ denote a normal and uniform
distribution, $(t_0', P')$ are the best-fit reference time and period
using only the pre-TESS transit times, and $(\sigma_{t_0'},
\sigma_{P'})$ are the corresponding uncertainties.

@@

% \section{Verifying the WASP-4 Archival Times}
% \label{sec:verify_archival_times}
% 
% \begin{figure*}[ht!]
%   \begin{center}
%     \leavevmode
%     \includegraphics[width=0.9\textwidth]{f7.pdf}
%   \end{center}
%   \caption{
%     {\bf Extra times for WASP-4b neither confirm nor refute the
%     	period-decay.}
%     Our timing analysis (Section~\ref{sec:timing}) imposes stringent
%     selection criteria for lightcurves.
%     Here we overplot additional points from the Exoplanet Transit
%     Database (ETD), and the discovery epoch from \citet{wilson_wasp-4b_2008}.
%     \label{fig:verify}
%   }
% \end{figure*}
% 
% The preceding analysis shows that the TESS time stamps are consistent
% at the requisite level with the ${\rm BJD}_{\rm TDB}$ reference.
% Another explanation for the WASP-4b timing variation could be
% incorrect archival times.
% As a sanity check on the times we included for our analysis in
% Table~1, we collected times from the Exoplanet Transit Database
% \citep[ETD; ][]{poddany_ETD_2010}.  We considered only the 12
% lightcurves of the highest self-reported quality (${\rm DQ} = 1$).  We
% then inspected each lightcurve by eye.  If the lightcurve showed
% substantial red noise, we discarded it.  This left 9 midtimes,
% reported in ${\rm HJD}_{\rm UTC}$. We converted the times to ${\rm
% BJD}_{\rm TDB}$ \citep{eastman_achieving_2010}.
% 
% The resulting times are shown in Figure \ref{fig:verify}.  Focusing on
% the epochs that overlap with \citet{huitson_gemini_2017}'s study, four
% of six possible times agree with the \citet{huitson_gemini_2017}
% times; the other two fall about 90 seconds early.  With their stated
% uncertainties, these 6 ETD data points are inconsistent with a
% constant period.  The 6 points came from 5 different observers.
% Systematic errors in absolute clock offsets by different observers
% should be uncorrelated; if there were such systematic errors in the
% ETD times, this could explain the many-sigma outliers.  Though the
% majority of the reported times do support the accuracy of the
% \citet{huitson_gemini_2017} times, a systematic procedure would call
% for averaging the ETD times.  If we did this, and discarded the four
% \citet{huitson_gemini_2017} epochs, it is likely that the evidence for
% period decay would become much weaker.  Since
% \citet{huitson_gemini_2017} documented their time system clearly, and
% produced lightcurves of extremely high quality, we included their times
% in our analysis.  From private correspondence with
% \citealt{huitson_gemini_2017}, the authors also stand by their
% time stamps. Since the ETD data come from
% heterogeneous sources, their time stamps are less clearly documented,
% and the times are thus more prone to systematic errors, we
% categorically omit them from consideration.
% 
% We also overplotted the discovery epoch reported by
% \citet{wilson_wasp-4b_2008}. 
% This epoch folded together WASP-S survey data, a partial FTS
% lightcurve, and a complete EulerCam lightcurve. 
% Though not explicitly stated by \citet{wilson_wasp-4b_2008}, the
% archival WASP-S lightcurves are in ${\rm HJD}_{\rm UTC}$, so they can
% be converted to ${\rm BJD}_{\rm TDB}$ without uncertainty in the
% absolute time reference (Collier-Cameron, priv.\ comm.).
% The discovery epoch falls about 2 minutes earlier than the epochs we
% used for the fit, and is visible in the lower left of Figure
% \ref{fig:verify}.
